R Notebook

### 2019-03-13 analysis/INLA/test-INLA-sensor.Rmd
AR1 spatial field running in INLA to predict temperature for all 6 years

```{r}
testdat <- tidy_survey_sets(raw, c("SYN HS","SYN QCS","SYN WCVI","SYN WCHG"), years = c(2015, 2016, 2017, 2018))
testdat <- gfplot:::interp_survey_bathymetry(testdat)
glimpse(testdat)
#View(testdat[[1]])
```

@@

```{r}
dat2 <- dat %>% filter(temperature_C_avg!="NA")

@@

  coords <- as.matrix(unique(d[, c("X", "Y")]))

    mesh <- INLA::inla.mesh.2d(
      offset = offset,
      boundary = bnd,
      max.edge = max_edge,
      cutoff = cutoff
    )
  }
  message("INLA max_edge = ", "c(", max_edge[[1]], ", ", max_edge[[2]], "), n = ", mesh$n)

@@

  #time_k <- dat$raw_time - min(dat$raw_time) + 1
  k <- max(time_k)
  time_lab <- paste0("Y", seq(1, k))
  dat[time_lab] <- 0
  dat[, time_lab[1]] <- 1
  for (j in seq_along(time_lab)) {
    dat[time_k == j, time_lab[j]] <- 1
  }
  #names(dat)[names(dat)=="Y1"] <- "intercept"

  # create an object for a Matérn model
  spde <- INLA::inla.spde2.matern(mesh = mesh, alpha = 3 / 2) #, # WHAT DOES ALPHS DO?????
  #                                prior.range = prior_range, 
  #                                prior.sigma = prior_sigma)

  # construct index for ar1 model ---- DON'T QUITE UNDERSTAND THIS
  iset <- INLA::inla.spde.make.index(
    name = "spatial.field",
    n.spde = mesh$n,
    n.group = k
  )

@@

 # projection matrix to map to observed points
 A <- INLA::inla.spde.make.A(
    mesh = mesh,
    loc = cbind(dat$xcoo, dat$ycoo),
    group = time_k
  )
  A.list <- list()
  A.list[[1]] <- A

@@

  # make projection points stack:
  sdat <- INLA::inla.stack(tag = "stdata",
                           data = list(y = dat$y),
                           A = A.list,
                           effects = effect.list
  )

  formula <- as.formula(paste(
    "y ~ -1 +",
    paste(covar_names, collapse = "+"),
    "+ f(spatial.field, model=spde)"
  ))

@@

# alternative to function, hard code

# samples = 10L
# pred_grid <- pg_one_year

@@

  # define empty arrays to store projections:
  latent_grid   <- array(0, dim = c(nrow(grid_locs), samples, length(yr_fe_indx)))
  spatial_field <- array(0, dim = c(nrow(grid_locs), samples, length(yr_fe_indx)))
  proj_matrix   <- INLA::inla.spde.make.A(mesh, loc = as.matrix(grid_locs))

>>>>

```{r}
testdat <- tidy_survey_sets(raw, c("SYN HS","SYN QCS","SYN WCVI","SYN WCHG"), years = c(2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018))
#testdat <- gfplot:::interp_survey_bathymetry(testdat)
glimpse(testdat)
#View(testdat[[1]])
```

@@

```{r}
dat2 <- dat %>% filter(temperature_C_avg!="NA") %>% filter(ssid==1)

@@

@@

  #time_k <- dat$raw_time - min(dat$raw_time) + 1
  k <- max(time_k)

  time_lab <- paste0("Y", seq(1, k))
  dat[time_lab] <- 0
  dat[, time_lab[1]] <- 1
  for (j in seq_along(time_lab)) {
    dat[time_k == j, time_lab[j]] <- 1
  }
  #names(dat)[names(dat)=="Y1"] <- "intercept"

  ## Select time knots and generate time mesh
  t_knots = seq(1, k, length = k)
  mesh1 = inla.mesh.1d(loc = t_knots, degree = 1, boundary = "free")

  ## Generate spatial mesh
  coords <- as.matrix(unique(d[, c("X", "Y")]))

    mesh2d <- INLA::inla.mesh.2d(
      offset = offset,
      boundary = bnd,
      max.edge = max_edge,
      cutoff = cutoff
    )
  }
  message("INLA max_edge = ", "c(", max_edge[[1]], ", ", max_edge[[2]], "), n = ", mesh$n)

  # create an object for a Matérn model
  spde <- INLA::inla.spde2.matern(mesh = mesh2d, alpha = 3 / 2) #, # WHAT DOES ALPHA DO?????
  #                                prior.range = prior_range, 
  #                                prior.sigma = prior_sigma)

  ## build the space time indices
  # index = inla.spde.make.index("space", n.spde = spde$n.spde, n.group = mesh1$m)
  # construct index for ar1 model ---- DON'T QUITE UNDERSTAND THIS
  iset <- INLA::inla.spde.make.index(
    name = "spatial.field",
    n.spde = spde$n.spde, 
    n.group = mesh1$m
  )

@@

 # projection matrix to map to observed points
 A <- INLA::inla.spde.make.A(
    mesh = mesh2d,
    loc = cbind(dat$xcoo, dat$ycoo),
    group = time_k, group.mesh = mesh1
  )
  A.list <- list()
  A.list[[1]] <- A

@@

  # make projection points stack:
  sdat <- INLA::inla.stack(tag = "stdata",
                           data = list(y = dat$y),
                           A = A.list,
                           effects = effect.list
  )
# 
#   formula <- as.formula(paste(
#     "y ~ -1 + ",
#     paste(covar_names, collapse = "+"), 
#     "+ f(spatial.field, model = spde, group = spatial.field.group, control.group = list(model = "ar1"))"))
#

  #covariates <- paste(covar_names, collapse = "+")

  formula <- y ~ -1 + depth_scaled+depth_scaled2+Y1+Y2+Y3+Y4+Y5+Y6 +
   f(spatial.field, model = spde, group = spatial.field.group, control.group = list(model = "ar1"))

 # formula <- as.formula(paste(
 #    "y ~ -1 + ",
 #    paste(covar_names, collapse = "+"), "+ 
 #      f(year_f, model = "rw2") +
 #      f(year_f2, depth_scaled, model = "rw2") +
 #      f(year_f3, depth_scaled2, model = "rw2") +
 #      f(spatial.field, model=spde)"
 #  ))
 #

@@

```{r}
# alternative to function, hard code
# samples = 10L
# pred_grid <- pg_one_year

@@

  # define empty arrays to store projections:
  latent_grid   <- array(0, dim = c(nrow(grid_locs), samples, length(yr_fe_indx)))
  spatial_field <- array(0, dim = c(nrow(grid_locs), samples, length(yr_fe_indx)))
  proj_matrix   <- INLA::inla.spde.make.A(mesh2d, loc = as.matrix(grid_locs))
